---
title: "eds221-day6-activities"
format: html
editor: visual
---

## Take a look at the data

```{r}
library(tidyverse)
library(here)
library(janitor)
```

```{r}
lobsters <- read_csv(here("data","Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>% 
  clean_names() %>% 
  uncount(count)

```

-   Create a summary table that finds the total counts (see: `n()`), and mean carapace lengths of lobsters observed in the dataset by **site** and **year**.

-   Create a ggplot graph of the number of total lobsters observed (y-axis) by year (x-axis) in the study, grouped (either aesthetically or by faceting) by site

```{r}
summary_lobsters <- lobsters |>
  group_by(site, year) |> 
  summarize(tot_counts =n() , mean_carapace = mean(size_mm))

```

Graph with aesthetic view

```{r}

lobster_plot <- ggplot(summary_lobsters, aes(x=year, y = tot_counts)) + 
  geom_point(aes(color = site)) + 
  labs(title = "Counts per Year")

lobster_plot
```

Graph with facet view

```{r}
lobster_plot <- ggplot(summary_lobsters, aes(x=year, y = tot_counts)) + 
  geom_point(aes(color = site)) + 
  labs(title = "Counts per Year") +
  facet_wrap(~site)

lobster_plot
```



-   Create a subset that only contains lobster data from 2020 (note: this should be from the original data you read in, not the summary table you created above)

```{r}

lobster_2020 <- subset(lobsters, year == 2020)
```

-   Write code (you can decide how to do this - there are a number of ways) to find the counts of lobsters observed at each site (only using site as the grouping factor) that are above and below the legal limit. **Hint:** You may want to add a new column `legal` that contains \"yes\" or \"no\" based on the size of the observed lobster (see `dplyr::case_when()` for a really nice way to do this), then use `group_by() %>% summarize(n())` or `dplyr::count()` to get counts by group within variables

```{r}

tot_lobster <- lobster_2020 |>
  mutate(legal= ifelse(size_mm > 79.76, "yes", "no")) |> 
  group_by(site, legal) |> 
  summarize(count = n())

```

Graph

```{r}
ggplot(tot_lobster, aes(fill =legal, x = site, y = count)) + 
  geom_col(position = "fill") 
```

### TASK 3

### **filter() practice**

1.  Create and store a subset that only contains lobsters from sites \"IVEE\", \"CARP\" and \"NAPL\". Check your output data frame to ensure that only those three sites exist.

2.  Create a subset that only contains lobsters observed in August.

3.  Create a subset with lobsters at Arroyo Quemado (AQUE) **OR** with a carapace length greater than 70 mm.

4.  Create a subset that does NOT include observations from Naples Reef (NAPL

```{r}

lobster_places <- lobsters |> 
  filter(site %in% c("IVEE", "AQUE", "CARP", "MOHK", "NAPL"))
lobster_august <- lobsters |> 
  filter(month == 8)
lobster_aque <- lobsters |> 
  filter(site == "AQUE" | size_mm > 70)
lobster_not_napl <- lobsters |> 
  filter(site != "NAPL")


```

### **group_by() %\>% summarize() practice**

5.  Find the mean and standard deviation of lobster carapace length, grouped by site

6.  Find the maximum carapace length by site and month.

```{r}

lobster_mean <- lobsters |> 
  group_by(site) |> 
  summarize(mean = mean(size_mm, na.rm = TRUE), std_dev = sd(size_mm, na.rm = TRUE))


lobster_max <- lobsters |> 
  group_by(site,month) |> 
  summarize(max_size_mm = max(size_mm, na.rm = TRUE))

```

### **mutate() practice**

7.  Add a new column that contains lobster carapace length converted to centimeters. Check output.

8.  Update the site column to all lowercase. Check output.

9.  Convert the area column to a character (not sure why you\'d want to do this, but try it anyway). Check output.

```{r}
lobster_cm <- lobsters |> 
  mutate(size_cm = size_mm/10)

lobster_lowersite <- lobsters |> 
  mutate(site = tolower(site))
```

### **case_when() practice**

10. Use `case_when()` to add a new column called `size_bin` that contains \"small\" if carapace size is \<= 70 mm, or \"large\" if it is greater than 70 mm. Check output.

11. Use `case_when()` to add a new column called `designation` that contains \"MPA\" if the site is \"IVEE\" or \"NAPL\", and \"not MPA\" for all other outcomes.

```{r}
# Question: How would you add na.rm = TRUE in size_bin?
lobster_large_small <- lobsters |> 
 mutate(
   size_bin = case_when(
     size_mm > 70 ~ "small",
     size_mm < 70 ~ "large",
     )
 )

loberster_mpa <- lobsters |> 
  mutate(
    destination = case_when(
      site %in% c("IVEE", "NAPL") ~ "MPA",
      site %in% c("AQUE", "CARP", "MOHK") ~ "not MPA"
    )
  )
```
